---
title: "EPİAŞ September Energy Dataset Analysis"
author: "Bilgen Yılmaz"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First Look at Dataset
After downloading the [dataset](https://rapor.epias.com.tr/rapor/xhtml/ptfSmfListeleme.xhtml) from [EPİAŞ official website](https://rapor.epias.com.tr/rapor/)
Understanding the columns and dataset process began.

Preparation was as follows, my table had non character values as column names due to language detection problem. So I changed the local language by using ` Sys.setlocale()` command.

```{r first, results='hide', message=FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)

Electricity_Data <- read.csv("C:/Users/cbilg/Documents/R/BDA503/Electricity_Assignment.csv")
Sys.setlocale(locale="Turkish_Turkey.1254")

```

To see the main elements of data, I used `glimpse()` command. As result I saw some of my values seem like they are chars instead of numeric. Also my time stamp was inproper. So I started to adjust to table to understand better.

```{r second, results='markup'}
glimpse(Electricity_Data)

```

## Creating Adjusted Table
While keeping old table, I created new table named `Electricity_Data_Adjusted` with the changes I wanted to make.

### Column names

I changed two of the column names to make it easier to read and use in codes.

```{r third, results='markup'}
Electricity_Data_Adjusted <- 
  Electricity_Data %>% rename(poz_deng_fiyat = Pozitif.Dengesizlik.Fiyatı..TL.MWh.) %>% 
  rename(neg_den_fiyat = Negatif.Dengesizlik.Fiyatı..TL.MWh.)

```

### Date formant

```{r fourth, results='markup'}
Electricity_Data_Adjusted$Tarih <- gsub(pattern = "\\.","-",Electricity_Data_Adjusted$Tarih)
Electricity_Data_Adjusted$Tarih <- as.POSIXct(Electricity_Data_Adjusted$Tarih,format = "%d-%m-%y %H:%M")

```

### Column types

```{r fifth, warning = FALSE}

Electricity_Data_Adjusted$PTF <- gsub(pattern = ",",".",Electricity_Data_Adjusted$PTF)
Electricity_Data_Adjusted$PTF <- as.double(Electricity_Data_Adjusted$PTF)

Electricity_Data_Adjusted$SMF <- gsub(pattern = ",",".",Electricity_Data_Adjusted$SMF)
Electricity_Data_Adjusted$SMF <- as.double(Electricity_Data_Adjusted$SMF)

Electricity_Data_Adjusted$poz_deng_fiyat <- gsub(pattern = ",",".",Electricity_Data_Adjusted$poz_deng_fiyat)
Electricity_Data_Adjusted$poz_deng_fiyat <- as.double(Electricity_Data_Adjusted$poz_deng_fiyat)

Electricity_Data_Adjusted$neg_den_fiyat <- gsub(pattern = ",",".",Electricity_Data_Adjusted$neg_den_fiyat)
Electricity_Data_Adjusted$neg_den_fiyat <- as.double(Electricity_Data_Adjusted$neg_den_fiyat)

```

### Summary of the final table

Finally, adjusted table's key metrics are as follows:

```{r sixth, results='markup'}
summary(Electricity_Data_Adjusted)

```

## Analysis of Dataset

As I looked into the table, energy need of summarized in three categorical value under the `SMF.Yön` column. `SMF.Yön`grouped summarized values were needed.
We can see when the energy imbalance is zero average PTF and SMF have the same value.

```{r seventh, results='markup', message=FALSE}
Electricity_Data_Adjusted %>%
  group_by(SMF.Yön) %>%
  summarize(tot_SMF = sum(SMF, na.rm=T), tot_PTF = sum(PTF), tot_PozDenge = sum(poz_deng_fiyat), 
            tot_NegDenge = sum(neg_den_fiyat, na.rm=T), avg_SMF =mean(SMF, na.rm=T), avg_PTF =mean(PTF))

```

### Examining the dataset daily basis
Using `Lubridate` library, I created daily total table for dataset.

```{r eighth, results='markup', message=FALSE}
Electricity_Data_Daily2 <- Electricity_Data_Adjusted %>%
  group_by(date = lubridate::date(Tarih)) %>%
  summarize(tot_SMF = sum(SMF, na.rm=T), tot_PTF = sum(PTF), tot_PozDenge = sum(poz_deng_fiyat), tot_NegDenge = sum(neg_den_fiyat),
            avg_SMF = mean(SMF, na.rm=T), avg_PTF = mean(PTF))

```

### Daily key metrics' visualization

This graph is showing us the daily average SMF value change over time. 
```{r nineth, results='markup', message=FALSE}
ggplot(Electricity_Data_Daily2, aes(x=date, y=avg_SMF))+
  geom_line()+
  labs(title="Daily Avg. SMF Graph",x ="Date", y = "Avg. SMF")  

```

This graph is showing us the daily average PTF value change over time. 
```{r tenth, results='markup', message=FALSE}
ggplot(Electricity_Data_Daily2, aes(x=date, y=avg_PTF))+
  geom_line()+
  labs(title="Daily Avg. PTF Graph",x ="Date", y = "Avg. PTF") 

```

### Analyzing daily values grouped by `SMF.Yön`

Daily values grouped by `SMF.Yön`and date.
```{r daily_data, results='markup', message=FALSE}
Electricity_Data_Daily <- Electricity_Data_Adjusted %>%
  group_by(date = lubridate::date(Tarih), SMF.Yön) %>%
  summarize(tot_SMF = sum(SMF), tot_PTF = sum(PTF), tot_PozDenge = sum(poz_deng_fiyat), tot_NegDenge = sum(neg_den_fiyat),
            avg_SMF = mean(SMF), avg_PTF = mean(PTF))

```

Bar plot of Energy Status on a daily basis to see their frequencies in September.
```{r daily_data_bar_graph, results='markup', message=FALSE}
ggplot(Electricity_Data_Daily, aes(x=SMF.Yön, fill=SMF.Yön))+
  geom_histogram(stat = "count")+
  labs(title="Electricity Status Histogram",x ="Status", y = "Total Count")

```

### Examining the dataset hourly basis

Examining the data set hourly basis:
```{r eleventh, results='markup', message=FALSE}
Electricity_Data_Hourly <- Electricity_Data_Adjusted %>%
  group_by(hour = lubridate::hour(Tarih)) %>%
  summarize(tot_SMF = sum(SMF, na.rm=T), tot_PTF = sum(PTF), tot_PozDenge = sum(poz_deng_fiyat), tot_NegDenge = sum(neg_den_fiyat),
            avg_SMF = mean(SMF, na.rm=T), avg_PTF = mean(PTF))

```

### Hourly key metrics' visualization

This graph is showing us the hourly average SMF value change over time. 
```{r hourly_smf_graph, results='markup', message=FALSE}
ggplot(Electricity_Data_Hourly, aes(x=hour, y=avg_SMF))+
  geom_line()+
  labs(title="Hourly Avg. SMF Graph",x ="Hour", y = "Avg. SMF")  

```

This graph is showing us the hourly average PTF value change over time. 
```{r hourly_ptf_graph, results='markup', message=FALSE}
ggplot(Electricity_Data_Hourly, aes(x=hour, y=avg_PTF))+
  geom_line()+
  labs(title="Hourly Avg. PTF Graph",x ="Hour", y = "Avg. PTF") 

```

This graph shows us comparison between hourly avg. SMF and hourly avg. PTF in a scatter plot: 
```{r smf_Vs_ptf, results='markup', message=FALSE}
ggplot(Electricity_Data_Hourly, aes(x=avg_SMF, y=avg_PTF))+
  geom_point()+
  labs(title="SMF vs. PTF Scatter",x ="Avg. Hourly SMF", y = "Avg. Hourly PTF") 

```



