---
title: "Diamonds Price Estimation"
author: "Bilgen Yılmaz"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse) 
library(rpart)
library(rpart.plot)
library(rattle)
library(ggplot2)
library(dplyr)

```

## Summary

Diamonds dataset consists of 10 columns and 53,940 rows. Colums are as follows;

  * `cut` : ordinal value of cut shape of the diamond. Values : Fair < Good < Very Good < Premium < Ideal
  * `carat` : numeric value of carats in a diamond
  * `clarity` : ordinal value of diamond's clarity.  Values : I1 < SI2 < SI1 < VS2 < VS1 < VVS2 < VVS1 < IF
  * `color` : ordinal value of diamond's color  Values : D < E < F < G < H < I < J
  * `depth` : numeric value of depth in a diamond
  * `table` : numeric value of diamonds flat surface on its facet
  * `price` : price of the diamond
  * `x` : numeric value
  * `y` : numeric value
  * `z` : numeric value
  

```{r diamonds}
summary(diamonds)
```

## Explanatory Data Analysis and Plots

We are expecting to find price value of an diamond by looking at its other attributes. Firstly we will examine columns' relation with price values in further plots.

### Plots for ordinal columns

In below tables, 'Cut', 'Clarity', 'Color' columns examined to see if there is a difference between factor in terms of 'Price'.

```{r plots1, figures-side, fig.show="hold", out.width="30%", echo=FALSE, message=FALSE, warning=FALSE}
diamonds %>% select(cut, price) %>% group_by(cut) %>% ggplot(aes(x=cut, y=price)) + geom_violin(scale='area') + labs(title =" Cut Plot by Price", x = "Cut", y = "Price") 
diamonds %>% select(clarity, price) %>% ggplot(aes(x=clarity, y=price)) + geom_violin(scale='area') + labs(title =" Clarity Plot by  Price", x = "Clarity", y = "Price") 
diamonds %>% select(color, price) %>% ggplot(aes(x=color, y=price)) + geom_violin(scale='area') + labs(title =" Color Plot by Price", x = "Color", y = "Price") 
```

### Plots for numeric columns

In below tables, 'Carat', 'Depth', 'X', 'Y', 'Z' columns examined to see if there is a relation/ trend with 'Price'

```{r plots2, figures-side, fig.show="hold", out.width="30%", echo=FALSE, message=FALSE, warning=FALSE}
diamonds %>% select(carat, price) %>% ggplot(aes(x=carat, y=price)) + geom_smooth(position='jitter') + geom_line(alpha= .2, aes(color = 'steelblue')) + labs(title =" Carat vs. Price", x = "Carat", y = "Price") 
diamonds %>% select(depth, price) %>% ggplot(aes(x=depth, y=price)) + geom_smooth(position='jitter') + geom_line(alpha= .2, aes(color='steelblue'))  + labs(title =" Depth vs. Price", x = "Depth", y = "Price") 
diamonds %>% select(x, price) %>% ggplot(aes(x= x, y=price)) + geom_point(aes(alpha=0.05)) +geom_smooth(position='jitter') + labs(title =" X vs. Price", x = "X", y = "Price") 
diamonds %>% select(y, price) %>% ggplot(aes(x= y, y=price)) + geom_point(aes(alpha=0.05)) +geom_smooth(position='jitter') + labs(title =" Y vs. Price", x = "Y", y = "Price")
diamonds %>% select(z, price) %>% ggplot(aes(x= z, y=price)) + geom_point(aes(alpha=0.05)) +geom_smooth(position='jitter') + labs(title =" Z vs. Price", x = "Z", y = "Price")
```


### Key Findings


* Cut, clarity, color has different scale of price frequency in them. Though it seem avg. price for types doesn't varies a lot.
* At some level, X, Y, Z attributes has a rising trend with price.
* Carat attribute has a rising trend with price.
* Depth doesn't seem to have an effect on price.


## PCA Analysis

To understand which column has higher correlation with price and to understand if we can do factor reduction, I did PCA analysis on data set. But because `cut`, `color`, `clarity` are not numerical, I transformed them into numerical values then applied PCA Analysis.

```{r numerictable, echo=FALSE, message=FALSE, warning=FALSE}
diamonds_numeric <- diamonds %>%
  transmute(color_n = as.numeric(diamonds$color),price,x,y,z, carat,
            cut_n = as.numeric(diamonds$cut),
            clarity_n =as.numeric(diamonds$clarity),
            color_n=as.numeric(diamonds$color))
# Correlation matrix of variables
cor(diamonds_numeric)
dia_pca_result3<-princomp(diamonds_numeric,cor=T)
# Summary PCA Result
summary(dia_pca_result3)

```

From correlation matrix, we can say 'X', 'Y', 'Z' and 'Carat' columns have the highest correlation with price. Meanwhile clarity has the highest negative correlation with price.
From PCA Analysis, we can say that with 5 component, we can explain %98.88 of the data.

## CART Model

### Model with 4 variable

According to PCA and correlation matrix. With 5 components I can explain %98.88 of the data. So I choose `carat`, `x`, `y`, `z` and `price` columns for the first model.

```{r cartmodel1, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(503)
diamonds_1 <- diamonds %>% select(carat, x,y,z,price)
diamonds_test1 <- diamonds_1 %>% mutate(diamond_id = row_number()) %>% 
  sample_frac(0.2) %>% ungroup()

diamonds_train1 <- anti_join(diamonds_1 %>% mutate(diamond_id = row_number()), 
                            diamonds_test1, by = "diamond_id")

diamonds_model1 <- rpart(price ~ ., data=diamonds_train1%>% select(-diamond_id))
fancyRpartPlot(diamonds_model1)

```

#### Sample Check

In `price_difference` column I have calculated the percentage of the difference between predicted price value and actual price.
In summary table, we can see we have a mean of %15 price difference. And between 1st and 3rd quartile we predicted price within -%15 and %38 range.

```{r modelcheck1, echo=FALSE, message=FALSE, warning=FALSE}

diamond_predict_1 <- predict(diamonds_model1)

in_sample_prediction <-
  cbind(
    diamond_predict_1 %>% tbl_df %>%
      transmute(price_predict = value),
    diamonds_train1 %>% tbl_df %>%
      transmute(price_actual = price)
  ) %>%
  mutate(price_difference = ((price_predict-price_actual)/(price_actual)))

summary(in_sample_prediction)

```


**Line plots of actual, predicted and difference columns of sample data**

```{r modelcheck12, figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}

ggplot(in_sample_prediction, aes(price_difference)) +
  geom_freqpoly(binwidth = 0.1)
ggplot(in_sample_prediction, aes(price_actual, color='blue')) +
  geom_freqpoly(binwidth = 250) +
  geom_freqpoly(aes(x=price_predict, color='green')) +
  labs(x = "Price")

```


#### Test Check

In `price_difference` column I have calculated the percentage of the difference between predicted price value and actual price.
Likewise sample summary table, we have a mean of %15 in `price_difference` column.

```{r testcheck1, echo=FALSE, message=FALSE, warning=FALSE}

diamond_predict_2 <- predict(diamonds_model1,newdata=diamonds_test1)

out_of_sample_prediction <-
  cbind(
    diamond_predict_2 %>% tbl_df %>%
      transmute(price_predict = value),
    diamonds_test1 %>% tbl_df %>%
      transmute(price_actual = price)
  ) %>%
  mutate(price_difference = ((price_predict/price_actual)-1))

summary(out_of_sample_prediction)

```

**Line plots of actual, predicted and difference columns of sample data**

```{r testcheck12, figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}

ggplot(out_of_sample_prediction, aes(price_difference)) +
  geom_freqpoly(binwidth = 0.1)

ggplot(out_of_sample_prediction, aes(price_actual, color='blue')) +
  geom_freqpoly(binwidth = 250) +
  geom_freqpoly(aes(x=price_predict, color='green'))+
  labs(x = "Price")

```


#### Differences


```{r differencecheck1, echo=FALSE, message=FALSE, warning=FALSE}
out_sample_prediction_sq <- out_of_sample_prediction %>% mutate(price_diff_value = price_predict-price_actual)
rmse1 <- sqrt(mean(out_sample_prediction_sq$price_diff_value^2)) 


in_sample_prediction_sq <- in_sample_prediction %>% mutate(price_diff_value = price_predict-price_actual)
rmse2 <- sqrt(mean(in_sample_prediction_sq$price_diff_value^2)) 

  
percent_sample1 <- in_sample_prediction %>% filter(between(price_difference,-0.20, 0.20)) %>% summarise(count=n()) / in_sample_prediction %>% summarise(count=n())
percent_test1 <- out_of_sample_prediction %>% filter(between(price_difference,-0.20, 0.20)) %>% summarise(count=n()) / out_of_sample_prediction %>% summarise(count=n())

```

To sum up, we can say we have mean price difference in value as `r rmse2` in sample and `r rmse1` in test data. %`r percent_sample1*100` of our predicted prices fell %20 above or %20 below of the actual prices in sample data. %`r percent_test1*100` of our predicted prices fell %20 above or %20 below of the actual prices in test data


### Model with all columns

Thus we saw the correlation matrix, with ordinal values converted to numeric values isn't a solid example of PCA Analysis. So I did CART with using all of the columns.

```{r cartmodel2, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(503)
diamonds_test <- diamonds %>% mutate(diamond_id = row_number()) %>% 
  group_by(cut, color, clarity) %>% sample_frac(0.2) %>% ungroup()

diamonds_train <- anti_join(diamonds %>% mutate(diamond_id = row_number()), 
                            diamonds_test, by = "diamond_id")
#running model

diamonds_model <- rpart(price ~ ., data=diamonds_train%>% select(-diamond_id))
fancyRpartPlot(diamonds_model)

```

As you can see from the table, With all of the columns `clarity`column made it its way to the decision criterias.

#### Sample Check

Sample and test data checked as the first CART model. With second CART, we have a difference mean of %14 and first quartile and third quartile are -%15, %35 respectively.

```{r modelcheck2, echo=FALSE, message=FALSE, warning=FALSE}

diamond_predict_3 <- predict(diamonds_model)

in_sample_prediction2 <-
  cbind(
    diamond_predict_3 %>% tbl_df %>%
      transmute(price_predict = value),
    diamonds_train %>% tbl_df %>%
      transmute(price_actual = price)
  ) %>%
  mutate(price_difference = ((price_predict-price_actual)/(price_actual)))

summary(in_sample_prediction2)

```

**Line plots of actual, predicted and difference columns of sample data**

```{r modelplot2, figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE }
ggplot(in_sample_prediction2, aes(price_difference)) +
  geom_freqpoly(binwidth = 0.1)

ggplot(in_sample_prediction2, aes(price_actual, color='blue')) +
  geom_freqpoly(binwidth = 250) +
  geom_freqpoly(aes(x=price_predict, color='green')) +
  labs(x = "Price")

```

#### Test Check

Close to sample check table. We have a mean of %14 in terms of price difference. And first quartile and third quartile are -%15 and %35 respectively.

```{r testcheck2, echo=FALSE, message=FALSE, warning=FALSE}
diamond_predict_4 <- predict(diamonds_model,newdata=diamonds_test)

out_of_sample_prediction2 <-
  cbind(
    diamond_predict_4 %>% tbl_df %>%
      transmute(price_predict = value),
    diamonds_test %>% tbl_df %>%
      transmute(price_actual = price)
  ) %>%
  mutate(price_difference = ((price_predict/price_actual)-1))

summary(out_of_sample_prediction2)

```

**Line plots of actual, predicted and difference columns of test data**

```{r testcheck13, figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ggplot(out_of_sample_prediction2, aes(price_difference)) +
  geom_freqpoly(binwidth = 0.1)

ggplot(out_of_sample_prediction2, aes(price_actual, color='blue')) +
  geom_freqpoly(binwidth = 250) +
  geom_freqpoly(aes(x=price_predict, color='green')) +
  labs(x = "Price")

```

```{r differencecheck2, echo=FALSE, message=FALSE, warning=FALSE}
out_sample_prediction_sq2 <- out_of_sample_prediction2 %>% mutate(price_diff_value = price_predict-price_actual)
rmse4 <- sqrt(mean(out_sample_prediction_sq2$price_diff_value^2)) 


in_sample_prediction_sq2 <- in_sample_prediction2 %>% mutate(price_diff_value = price_predict-price_actual)
rmse3 <- sqrt(mean(in_sample_prediction_sq2$price_diff_value^2)) 

  
percent_sample2 <- in_sample_prediction2 %>% filter(between(price_difference,-0.20, 0.20)) %>% summarise(count=n()) / in_sample_prediction %>% summarise(count=n())
percent_test2 <- out_of_sample_prediction2 %>% filter(between(price_difference,-0.20, 0.20)) %>% summarise(count=n()) / out_of_sample_prediction %>% summarise(count=n())

```

We can say we have mean price difference in value as `r rmse3` in sample and `r rmse4` in test data
%`r percent_sample2 * 100` of our predicted prices fell %20 above or %20 below of the actual prices in sample data
%`r percent_test2 * 100` of our predicted prices fell %20 above or %20 below of the actual prices in test data


Differences in values and the total number of the predicted values fall between -%20 to %20 from actual values was higher in the second CART. So I choose to go with all the columns in data table.

#### References

http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/114-mca-multiple-correspondence-analysis-in-r-essentials/
https://alice86.github.io/2018/04/08/Factor-Analysis-on-Ordinal-Data-example-in-R-(psych,-homals)/
https://rpubs.com/DocOfi/342740
https://mef-bda503.github.io/archive/fall17/files/intro_to_ml.html#properties
https://mef-bda503.github.io/archive/fall17/files/intro_to_ml_2.html
